% This file defines the propert to be inserted into vespers.tex  In this way
% multiple feasts can be typeset very quickly.  vespers.tex should not normally
% need changing.  Note that this is not the most readable way to insert text
% into a LaTeX document, but it is the most powerful: the macros defined here
% are directly excecuted when building the document.

% For the title page
\newcommand{\feast}{S. Joannis Bosco Confessoris}

% Translations of antiphons
\newcommand{\antonetranslation}{{O Lord, thou has searched me out and
    knowest me.}}
\newcommand{\anttwotranslation}{{Wonderful are thy works, and
    that my soul knoweth right well.}}
\newcommand{\antthreetranslation}{{Forsake not thou me O Lord,
    the strength of my salvation.}}
\newcommand{\antfourtranslation}{{I have cried to thee O Lord:
    hear me.}}
\newcommand{\antfivetranslation}{{Bring my soul out of prison, O
    Lord, that I may praise thy name.}}

% Text of psalms, formatted.  This can be taken directly from the formatted
% psalms provided.  Translations are after the psalm to which the correspond to
% try to avoid mistakes.
\newcommand{\psalmOne}{
  Intellexísti cogitatiónes \textbf{me}as de \textbf{lon}ge:~\*\\
  \vin sémitam meam, et funículum meum in\textbf{ves}ti\textbf{gás}ti.\\
  Et omnes vias meas \textbf{præ}vi\textbf{dís}ti:~\*\\
  \vin quia non est sermo in \textbf{lin}gua \textbf{me}a.\\
  Ecce, Dómine, tu cognovísti ómnia novíssima, \textbf{et} an\textbf{tí}qua:~\*\\
  \vin tu formásti me, et posuísti super me \textbf{ma}num \textbf{tu}am.\\
  Mirábilis facta est sciéntia \textbf{tu}a \textbf{ex} me:~\*\\
  \vin confortáta est, et non póte\textbf{ro} ad \textbf{e}am.\\
  Quo ibo a \textbf{spí}ritu \textbf{tu}o?~\*\\
  \vin et quo a fácie \textbf{tu}a \textbf{fú}giam?\\
  Si ascéndero in \textbf{cæ}lum, tu \textbf{illic} es:~\*\\
  \vin si descéndero in in\textbf{fér}num, \textbf{ad}es.\\
  Si súmpsero pennas \textbf{me}as di\textbf{lúcu}lo,~\*\\
  \vin et habitávero in ex\textbf{tré}mis \textbf{ma}ris.\\
  Etenim illuc manus \textbf{tu}a de\textbf{dúcet} me:~\*\\
  \vin et tenébit me \textbf{déx}tera \textbf{tu}a.\\
  Et dixi: Fórsitan ténebræ \textbf{con}cul\textbf{cábunt} me:~\*\\
  \vin et nox illuminátio mea in de\textbf{lí}ciis \textbf{me}is.\\
  Quia ténebræ non obscurabúntur a te,~\+\\
  \vin et nox sicut dies il\textbf{lu}mi\textbf{nábi}tur:~\*\\
  \vin sicut ténebræ ejus, ita et \textbf{lu}men \textbf{e}jus.\\
  Quia tu possedísti \textbf{re}nes \textbf{me}os:~\*\\
  \vin suscepísti me de útero \textbf{ma}tris \textbf{me}æ.\\
  Glória \textbf{Pa}tri, et \textbf{Fíli}o,~\*\\
  \vin et Spi\textbf{rí}tui \textbf{Sanc}to.\\
  Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~\*\\
  \vin et in sǽcula sæcu\textbf{ló}rum. \textbf{A}men.\\
}

\newcommand{\translationOne}{
  Lord, I lie open to thy scrutiny; thou knowest me,  knowest when
  sit down and when I rise up again, canst read my thoughts from far
  away.  Walk I or sleep I, thou canst tell; no movement of mine but
  thou art watching it.  Before ever the words are framed on my lips,
  all my thought is known to thee;  rearguard and vanguard, thou dost
  compass me about, thy hand still laid upon me.  Such wisdom as thine
  is far beyond my reach, no thought of mine can attain it.

  Where can I go, then, to take refuge from thy spirit, to hide from
  thy view?  If I should climb up to heaven, thou art there; if I sink
  down to the world beneath, thou art present still.  If I could wing
  my way eastwards, or find a dwelling beyond the western sea,  still
  would I find thee beckoning to me, thy right hand upholding me.  Or
  perhaps I would think to bury myself in darkness; night should
  surround me, friendlier than day;  but no, darkness is no
  hiding-place from thee, with thee the night shines clear as day
  itself; light and dark are one.

  Author, thou, of my inmost being, didst thou not form me in my
  mother’s womb?}


\newcommand{\psalmTwo}{
  Non est occultátum os meum a te, quod fecísti in \emph{oc}\textbf{cúl}to:~\*\\
  \vin et substántia mea in inferió\emph{ribus} \textbf{ter}ræ.\\
  Imperféctum meum vidérunt óculi tui,~\+\\
  \vin et in libro tuo omnes \emph{scri}\textbf{bén}tur:~\*\\
  \vin dies formabúntur, et ne\emph{mo} \emph{in} \textbf{e}is.\\
  Mihi autem nimis honorificáti sunt amíci tu\emph{i}, \textbf{De}us:~\*\\
  \vin nimis confortátus est principá\emph{tus} \emph{e}\textbf{ó}rum.\\
  Dinumerábo eos, et super arénam multipli\emph{ca}\textbf{bún}tur:~\*\\
  \vin exsurréxi, et ad\emph{huc} \emph{sum} \textbf{te}cum.\\
  Si occíderis, Deus, pec\emph{ca}\textbf{tó}res:~\*\\
  \vin viri sánguinum, decli\emph{náte} \textbf{a} me.\\
  Quia dícitis in cogita\emph{ti}\textbf{ó}ne:~\*\\
  \vin Accípient in vanitáte civi\emph{tátes} \textbf{tu}as.\\
  Nonne qui odérunt te, Dómi\emph{ne}, \textbf{ó}deram?~\*\\
  \vin et super inimícos tuos \emph{tabe}\textbf{scé}bam?\\
  Perfécto ódio óde\emph{ram} \textbf{il}los:~\*\\
  \vin et inimíci fac\emph{ti} \emph{sunt} \textbf{mi}hi.\\
  Proba me, Deus, et scito \emph{cor} \textbf{me}um:~\*\\
  \vin intérroga me, et cognósce sé\emph{mitas} \textbf{me}as.\\
  Et vide, si via iniquitátis \emph{in} \textbf{me} est:~\*\\
  \vin et deduc me in vi\emph{a} \emph{æ}\textbf{tér}na.\\
  Glória Patri, \emph{et} \textbf{Fí}lio,~\*\\
  \vin et Spirí\emph{tui} \textbf{Sanc}to.\\
  Sicut erat in princípio, et nunc, \emph{et} \textbf{sem}per,~\*\\
  \vin et in sǽcula sæcu\emph{lórum}. \textbf{A}men.\\
}

\newcommand{\translationTwo}{  I praise thee for my wondrous fashioning, for all
  the wonders of thy creation. Of my soul thou hast full knowledge, 
  and this mortal frame had no mysteries for thee, who didst contrive it
  in secret, devise its pattern, there in the dark recesses of the
  earth.  All my acts thy eyes have seen, all are set down already in
  thy record; my days were numbered before ever they came to be.

  A riddle, O my God, thy dealings with me, so vast their scope! 
  As well count the sand, as try to fathom them; and, were that skill
  mine, thy own being still confronts me.  O God, wouldst thou but
  make an end of the wicked! Murderers, keep your distance from me! 
  Treacherously they rebel against thee, faithlessly set thee at
  defiance.  Lord, do I not hate the men who hate thee, am I not sick
  at heart over their rebellion?  Surpassing hatred I bear them, count
  them my sworn enemies.  Scrutinize me, O God, as thou wilt, and read
  my heart; put me to the test, and examine my restless thoughts.  See
  if on any false paths my heart is set, and thyself lead me in the ways
  of old.
}

\newcommand{\psalmThree}{
  Qui cogitavérunt iniquitá\emph{tes} \emph{in} \textbf{cor}de:~\*\\
  \vin tota die constituébant \textbf{prǽ}lia.\\
  Acuérunt linguas suas sic\emph{ut} \emph{ser}\textbf{pén}tis:~\*\\
  \vin venénum áspidum sub lábiis e\textbf{ó}rum.\\
  Custódi me, Dómine, de manu \emph{pecca}\textbf{tó}ris:~\*\\
  \vin et ab homínibus iníquis éri\textbf{pe} me.\\
  Qui cogitavérunt supplantáre \emph{gressus} \textbf{me}os:~\*\\
  \vin abscondérunt supérbi láqueum \textbf{mi}hi:\\
  Et funes extendé\emph{runt} \emph{in} \textbf{lá}queum:~\*\\
  \vin juxta iter scándalum posuérunt \textbf{mi}hi.\\
  Dixi Dómino: Deus \emph{meus} \textbf{es} tu:~\*\\
  \vin exáudi, Dómine, vocem deprecatiónis \textbf{me}æ.\\
  Dómine, Dómine, virtus sa\emph{lútis} \textbf{me}æ:~\*\\
  \vin obumbrásti super caput meum in die \textbf{bel}li.\\
  Ne tradas me, Dómine, a desidério meo peccatóri:~\+\\
  \vin cogita\emph{vérunt} \textbf{con}tra me,~\*\\
  \vin ne derelínquas me, ne forte exal\textbf{tén}tur.\\
  Caput circúi\emph{tus} \emph{e}\textbf{ó}rum:~\*\\
  \vin labor labiórum ipsórum opériet \textbf{e}os.\\
  Cadent super eos carbónes,~\+\\
  \vin in ignem dejí\emph{cies} \textbf{e}os:~\*\\
  \vin in misériis non sub\textbf{sís}tent.\\
  Vir linguósus non dirigé\emph{tur} \emph{in} \textbf{ter}ra:~\*\\
  \vin virum injústum mala cápient in in\textbf{tér}itu.\\
  Cognóvi quia fáciet Dóminus judí\emph{cium} \textbf{ín}opis:~\*\\
  \vin et vindíctam \textbf{páu}perum.\\
  Verúmtamen justi confitebúntur nó\emph{mini} \textbf{tu}o:~\*\\
  \vin et habitábunt recti cum vultu \textbf{tu}o.\\
  Glória Pa\emph{tri}, \emph{et} \textbf{Fí}lio,~\*\\
  \vin et Spirítui \textbf{Sanc}to.\\
  Sicut erat in princípio, et \emph{nunc}, \emph{et} \textbf{sem}per,~\*\\
  \vin et in sǽcula sæculórum. \textbf{A}men.\\
}

\newcommand{\translationThree}{
  Rescue me, Lord, from human malice, save me from the lovers of
  oppression,  always plotting treachery in their hearts, always at
  their quarrelling,  tongues sharp as the tongues of serpents, lips
  that hide the poison of adders.  Preserve me Lord, from the power of
  sinful men, save me from these lovers of oppression who are plotting
  to trip my feet.  What hidden snares they set for me, these tyrants,
  what nets they spread to catch me, what traps they lay in my path!

  To the Lord I make my appeal, Thou art my God, listen, Lord, to the
  voice that pleads with thee.  My Lord, my Master, my strong
  deliverer, it is thou that shieldest my head in the day of battle. 
  Lord, do not let malice have its way with me, do not prosper its evil
  designs.  They carry their heads high as they close in around me;
  let their conspiracy prove its own undoing;  let burning coals rain
  down on them, be they cast into a pit whence they shall rise no
  more.  Not long the blasphemer’s time on earth: misfortune will
  overtake the oppressor unawares.  Can I doubt that the Lord will
  avenge the helpless, will grant the poor redress?  Honest men will
  yet live to praise thy name; upright hearts to enjoy the smile of thy
  favour.
}  

\newcommand{\psalmFour}{
  Dirigátur orátio mea sicut incénsum in conspéctu \textbf{tu}o:~\*\\
  \vin elevátio mánuum meárum sacrifícium \emph{vesper}\textbf{tí}num.\\
  Pone, Dómine, custódiam ori \textbf{me}o:~\*\\
  \vin et óstium circumstántiæ lá\emph{biis} \textbf{me}is.\\
  Non declínes cor meum in verba ma\textbf{lí}tiæ:~\*\\
  \vin ad excusándas excusatiónes \emph{in} \emph{pec}\textbf{cá}tis.\\
  Cum homínibus operántibus iniqui\textbf{tá}tem:~\*\\
  \vin et non communicábo cum eléc\emph{tis} \emph{e}\textbf{ó}rum.\\
  Corrípiet me justus in misericórdia, et incre\textbf{pá}bit me:~\*\\
  \vin óleum autem peccatóris non impínguet \emph{caput} \textbf{me}um.\\
  Quóniam adhuc et orátio mea in beneplácitis e\textbf{ó}rum:~\*\\
  \vin absórpti sunt juncti petræ júdi\emph{ces} \emph{e}\textbf{ó}rum.\\
  Audient verba mea quóniam potu\textbf{é}runt:~\*\\
  \vin sicut crassitúdo terræ erúpta est \emph{super} \textbf{ter}ram.\\
  Dissipáta sunt ossa nostra secus inférnum:~\+\\
  \vin quia ad te, Dómine, Dómine, óculi \textbf{me}i:~\*\\
  \vin in te sperávi, non áuferas á\emph{nimam} \textbf{me}am.\\
  Custódi me a láqueo, quem statuérunt \textbf{mi}hi:~\*\\
  \vin et a scándalis operántium in\emph{iqui}\textbf{tá}tem.\\
  Glória Patri, et \textbf{Fí}lio,~\*\\
  \vin et Spirí\emph{tui} \textbf{Sanc}to.\\
  Sicut erat in princípio, et nunc, et \textbf{sem}per,~\*\\
  \vin et in sǽcula sæcu\emph{lórum}. \textbf{A}men.\\
}

\newcommand{\translationFour}{
  Come quickly, Lord, at my cry for succour; do not let my appeal to
  thee go unheard.  Welcome as incense-smoke let my prayer rise up
  before thee; when I lift up my hands, be it acceptable as the evening
  sacrifice.  Lord, set a guard on my mouth, post a sentry before my
  lips;  do not turn my heart towards thoughts of evil, and deeds of
  treachery; never let me take part with the wrong-doers, and share the
  banquet with them.  Rather let some just man deal me heavy blows;
  this shall be his kindness to me; reprove me, and it shall be balm
  poured over me; such unction never will this head refuse. Their
  injuries I will still greet with a prayer\ldots

  \ldots My words have won their hearts, a people that had seen their
  chieftains hurled down the rock-face,  a people whose bones lie
  scattered at the grave’s mouth, like seed when the earth is cloven
  into furrows.

  To thee these eyes look, my Lord, my Master; in thee I trust; let
  not my life be forfeit.  Preserve me from the ambush they have laid
  for me, from the snares of the wrong-doers.  Into their own net,
  sinner upon sinner, may they fall, and I pass on in safety.}

\newcommand{\psalmFive}{
  Effúndo in conspéctu ejus orati\textbf{ó}nem \textbf{me}am,~\*\\
  \vin et tribulatiónem meam ante \textbf{ip}sum pro\textbf{nún}tio.\\
  In deficiéndo ex me \textbf{spí}ritum \textbf{me}um:~\*\\
  \vin et tu cognovísti \textbf{sé}mitas \textbf{me}as.\\
  In via hac, qua \textbf{am}bu\textbf{lá}bam,~\*\\
  \vin abscondérunt \textbf{lá}queum \textbf{mi}hi.\\
  Considerábam ad déxteram, \textbf{et} vi\textbf{dé}bam:~\*\\
  \vin et non erat qui co\textbf{gnó}sce\textbf{ret} me.\\
  Périit \textbf{fu}ga \textbf{a} me:~\*\\
  \vin et non est qui requírat \textbf{á}nimam \textbf{me}am.\\
  Clamávi ad te, Dómine,~\+\\
  \vin dixi: Tu \textbf{es} spes \textbf{me}a,~\*\\
  \vin pórtio mea in \textbf{ter}ra vi\textbf{vén}tium.\\
  Inténde ad deprecati\textbf{ó}nem \textbf{me}am:~\*\\
  \vin quia humili\textbf{á}tus sum \textbf{ni}mis.\\
  Líbera me a perse\textbf{quén}ti\textbf{bus} me:~\*\\
  \vin quia confor\textbf{tá}ti sunt \textbf{su}per me.\\
  Educ de custódia ánimam meam ad confiténdum \textbf{nó}mini \textbf{tu}o:~\*\\
  \vin me exspéctant justi, donec re\textbf{trí}buas \textbf{mi}hi.\\
  Glória \textbf{Pa}tri, et \textbf{Fíli}o,~\*\\
  \vin et Spi\textbf{rí}tui \textbf{Sanc}to.\\
  Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~\*\\
  \vin et in sǽcula sæcu\textbf{ló}rum. \textbf{A}men.\\
}

\newcommand{\translationFive}{  Loud is my cry to the Lord, the
  prayer I utter for the Lord’s mercy,  as I pour out my complaint
  before him, tell him of the affliction I endure.  My heart is ready
  to faint within me, but thou art watching over my path. They lie in
  ambush for me, there by the wayside;  I look to the right of me, and
  find none to take my part; all hope of escape is cut off from me, none
  is concerned for my safety.  To thee, Lord, I cry, claiming thee for
  my only refuge, all that is left me in this world of living men. 
  Listen, then, to my plea; thou seest me all defenceless. Rescue me
  from persecutors who are too strong for me; restore liberty to a
  captive soul.  What thanks, then, will I give to thy name, honest
  hearts all about me, rejoicing to see thy favour restored!  }

% The Chapter will be inserted directly in a multicol environment, so we just
% provide the latin (\l{}) and english (\e{})
\newcommand{\capitulum}{ \l{Beátus vir, qui invéntus est sine mácula,
    et qui post aurum non ábiit, nec sperávit in pecúnia et
    thesáuris. Quis est hic, et laudábimus eum? fecit enim mirabília
    in vita sua.}
  \e{Blessed is the man that is found without blemish: and that hath not
    gone after gold, nor put his trust in money nor in treasures. Who is
    he, and we will praise him? for he hath done wonderful things in his
    life.}}

% Hymn
\newcommand{\hymntranslation}{
  This the confessor of the Lord, whose triumph\\
  Now all the faithful celebrate, with gladness\\
  On this feast day year by year receiveth\\
  Merited honors.

  Saintly and prudent, modest in behaviour,\\
  Peaceful and sober, chaste was he, and lowly.\\
  While that life’s vigour, coursing through his members,\\
  Quickened his being.

  Sick ones of old time, to his tomb resorting,\\
  Sorely by ailments manifold afflicted,\\
  Oft-times have welcomed health and strength returning,\\
  At his petition.

  Whence we in chorus gladly do him honour,\\
  Chanting his praises with devout affection,\\
  That in his merits we may have a portion,\\
  Now and forever.

  His be the glory, power and salvation,\\
  Who over all things reigneth in the highest,\\
  Earth’s mighty fabric ruling and directing,\\
  Onely and Trinal.
  Amen.}

% The hymn is in a multicol environment
\newcommand{\versus}{
  \l{\v~Iustum dedúxit Dóminus per vias rectas.}
  \e{\v~The Lord has lead the just in right paths.}
  \l{\r~Et osténdit illi regnum Dei.}
  \e{\r~And showed unto him the kingdom of God.}
}

% Magnificat Antiphon
\newcommand{\magnificatAntiphonTranslation}{Lo, a servant of God who
  esteemed but little things earthly. And by word and work laid him up
  treasure in heaven.}

% Collect
\newcommand{\oratio}{
  \l{Deus, qui sanctum Ioánnem Confessórem tuum
    adolescéntium patrem et magístrum excitásti, ac per eum,
    auxiliatríce Vírgine María, novas in Ecclésia tua famílias
    floréscere voluísti: concéde, quǽsumus; ut eódem caritátis igne
    succénsi, ánimas quǽrere, tibíque soli servíre valeámus.}
  \e{O God, who didst raise up thy blessed Confessor John, to be a
    father and teacher of youth, and didst will that through him, with
    the help of the Virgin Mary, new families should flourish in the
    Church, grant we pray that enkindled with the same fire of charity
    we may be strong to seek after souls and to serve thee alone.}
  \l{Per Dóminum nostrum Iesum Christum, Fílium tuum: qui tecum vivit
    et regnat in unitáte Spíritus Sancti Deus, per ómnia sǽcula
    sæculórum.}  \e{Through Jesus Christ, thy Son our Lord, Who liveth
    and reigneth with thee, in the unity of the Holy Ghost, ever one
    God, world without end.}  \amen }

% Commemorations if any.  If empty this will be skipped.  It should contain *all
% the text* of the commemorations, including responses.
\newcommand{\com}{}             %no commemoratio


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "vespers"
%%% End:
