% This file defines the propert to be inserted into vespers.tex  In this way
% multiple feasts can be typeset very quickly.  vespers.tex should not normally
% need changing.  Note that this is not the most readable way to insert text
% into a LaTeX document, but it is the most powerful: the macros defined here
% are directly excecuted when building the document.

% For the title page
\newcommand{\feast}{S. Joseph Opificis}

\newcommand{\lenten}{} % set to true in lent.

% Translations of antiphons
\newcommand{\antonetranslation}{God whose working fashioned the world, has
  stationed man to work and till the earth, alleluia.}
\newcommand{\anttwotranslation}{Christ, the Son of God, * deigned to work with his hands, alleluia.}

% Define these if required.  They will be ignored if anttwotranslation is blank.
\newcommand{\antthreetranslation}{The craftsman, * holy Joseph, faithfully exercising his trade, shines forth as a marvellous example of work, alleluia.}
\newcommand{\antfourtranslation}{Faithful servant, * and prudent, whom the Lord hath appointed over his family, alleluia.}
\newcommand{\antfivetranslation}{O Joseph, holy workman, * protect our work, alleluia.}


% Text of psalms, formatted. This can be taken directly from the formatted
% psalms provided. Alternatively you can use parsing script to generate a
% psalms.tex and include it here. For more information, see the documentation.
\input{psalms.tex}

% \newcommand{\psalmOne}
% \newcommand{\psalmTwo}
% \newcommand{\psalmThree}
% \newcommand{\psalmFour}
% \newcommand{\psalmFive}
% \newcommand{\magnificat}

\newcommand{\translationOne}{
  To the Master I serve the Lord’s promise was given, Sit here at my right hand while I make thy enemies a footstool under thy feet.
  The Lord will make thy empire spring up like a branch out of Sion; thou art to bear rule in the midst of thy enemies.
  From birth, princely state shall be thine, holy and glorious; thou art my son, born like dew before the day-star rises.
  The Lord has sworn an oath there is no retracting, Thou art a priest for ever in the line of Melchisedech.
  At thy right hand, the Lord will beat down kings in the day of his vengeance;
  he will pass sentence on the nations, heap high the bodies, scatter far and wide the heads of the slain.
  Let him but drink of the brook by the wayside, he will lift up his head in victory.
}


\newcommand{\translationTwo}{
  \textbf{A}ll my heart goes out to the Lord in praise, \textbf{B}efore the assembly where the just are gathered.
  \textbf{C}hant we the Lord’s wondrous doings, \textbf{d}elight and study of all who love him.
  \textbf{E}ver his deeds are high and glorious, \textbf{f}aithful he abides to all eternity.
  \textbf{G}reat deeds, that he keeps still in remembrance!
  \textbf{H}e, the Lord, is kind and merciful. \textbf{I}n abundance he fed the men who feared him, \textbf{k}eeping his covenant for ever.
  \textbf{L}ordly the power he shewed his people,
  \textbf{m}aking the lands of the heathen their possession. \textbf{N}o act but shews him just and faithful; \textbf{o}f his decrees there is no relenting.
  \textbf{P}erpetual time shall leave them changeless; \textbf{r}ight and truth are their foundation.
  \textbf{S}o he has brought our race deliverance; \textbf{t}o all eternity stands his covenant. \textbf{U}nutterable is his name and worshipful;
   \textbf{v}ain without his fear is learning. \textbf{W}ise evermore are you who follow it; \textbf{y}ours the prize that lasts for ever.
}

\newcommand{\translationThree}{
  \textbf{A} blessed man is he, who fears the Lord, \textbf{b}earing great love to his commandments.
  \textbf{C}hildren of his shall win renown in their country; \textbf{d}o right, and thy sons shall find a blessing.
  \textbf{E}ase shall dwell in his house, and great prosperity; \textbf{f}ame shall ever record his bounty.
  \textbf{G}ood men see a light dawn in darkness; \textbf{h}is light, who is merciful, kind and faithful.
  \textbf{I}t goes well with the man who lends in pity, \textbf{j}ust and merciful in his dealings.
  \textbf{L}ength of days shall leave him still unshaken; \textbf{m}en will remember the just for ever.
  \textbf{N}o fear shall he have of evil tidings; \textbf{o}n the Lord his hope is fixed unchangeably.
  \textbf{P}atient his heart remains and stedfast, \textbf{q}uietly he waits for the downfall of his enemies.
  \textbf{R}ich are his alms to the needy; \textbf{s}till his bounty abides in memory. \textbf{T}he Lord will lift up his head in triumph;
   \textbf{u}ngodly men are ill content to see it. \textbf{V}ainly they gnash their teeth in envy; \textbf{w}orldly hopes must fade and perish.

}

\newcommand{\translationFour}{
  Praise the Lord, you that are his servants, praise the name of the Lord together.
  Blessed be the Lord’s name at all times, from this day to all eternity;
  from the sun’s rise to the sun’s setting let the Lord’s name be praised continually.
  The Lord is sovereign king of all the nations; his glory is high above the heavens.
  Who is like the Lord our God, so high above us,
  that stoops to regard both heaven and earth,
  lifting up the poor from the dust he lay in, raising the beggar out of his dung-hill,
  to find him a place among the princes, the princes that rule over his people?
  He gives the barren woman a home to dwell in, a mother rejoicing in her children.
}

\newcommand{\translationFive}{
  Praise the Lord, all you Gentiles, let all the nations of the world do him honour.
  Abundant has his mercy been towards us; the Lord remains faithful to his word for ever.
}

% The Chapter will be inserted directly in a multicol environment, so we just
% provide the latin (\l{}) and english (\e{})
\newcommand{\capitulum}{
  \l{Fratres: Caritátem habéte, quod est vínculum perfectiónis, et pax Christi exsúltet in córdibus vestris, et grati estóte.}
  \e{Brethren: have charity, which is the bond of perfection, and let the peace of Christ rejoice in your hearts, and be ye thankful.}
}

% Hymn
\newcommand{\hymntranslation}{
  Let angels chant thy praise, pure spouse of purest bride,\\
  While Christendom's sweet choirs the gladsome strains repeat,\\
  To tell thy wondrous fame, to raise the pealing hymn,\\
  Wherewith we all thy glory greet.

  When doubts and bitter fears thy heavy heart oppressed,\\
  And filled thy righteous soul with sorrow and dismay,\\
  An angel quickly came, the wondrous secret told,\\
  And drove thy anxious griefs away.

  Thy arms thy new-born Lord, with tender joy embrace;\\
  Him then to Egypt's land thy watchful care doth bring;\\
  Him in the temple's courts once lost thou dost regain,\\
  And 'mid thy tears dost greet thy King.

  Not till death's pangs are o'er do others gain their crown,\\
  But, Joseph, unto thee the blessed lot was given\\
  While life did yet endure, thy God to see and know,\\
  As do the saints above in heaven.

  Grant us, great Trinity, for Joseph's holy sake,\\
  In highest bliss and love, above the stars to reign,\\
  That we in joy with him may praise our loving God,\\
  And sing our glad eternal strain.
  Amen.
}

\newcommand{\versus}{
  \l{\v~Ora pro nobis, sancte Ioseph, allelúia.}
  \e{\v~Pray for us, holy Joseph, alleluia.}
  \l{\r~Labórum nostrórum protéctor fidélis, allelúia.}
  \e{\r~Faithful protector of our labours, alleluia.}
}

% Magnificat Antiphon
\newcommand{\magnificatAntiphonTranslation}{And Jesus himself * was beginning about the age of thirty years; being (as it was supposed) the son of Joseph, alleluia.}

% Collect
\newcommand{\oratio}{
  \l{Rerum cónditor Deus, qui legem labóris humáno géneri statuísti: concéde propítius; ut, sancti Ioseph exémplo et patrocínio, ópera perficiámus quae prǽcipis, et prǽmia consequámur quæ promíttis.}
  \e{God, Creator of all things, who didst lay on the human race the law of labour: graciously grant; that by following the example of Saint Joseph and under his patronage, we may carry out the work thou dost command, and obtain the reward thou dost promise.}
  \l{Per Dóminum nostrum Iesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.}
  \e{Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.}
  \amen
}

% Commemorations if any.  If empty this will be skipped.  It should contain *all
% the text* of the commemorations, including responses.
\newcommand{\com}{
  \red{Commemoratio Dominica II Post Pascha}
  \black{
    \l{\ant Ego sum pastor bonus, qui pasco oves meas, et pro óvibus meis pono ánimam meam, allelúia.}
    \e{I am the Good Shepherd, Who feed My sheep: and I lay down My life for My sheep. Alleluia.}
    \l{\v~Mane nobíscum, Dómine, allelúia.}
    \e{\v~Stay with us, O Lord, alleluia.}
    \l{\r~Quóniam advesperáscit, allelúia.}
    \e{\r~For the evening falls, alleluia.}
    \oremus
    \l{Deus, qui in Fílii tui humilitáte iacéntem mundum erexísti: fidélibus tuis perpétuam concéde lætítiam; ut, quos perpétuæ mortis eripuísti cásibus, gáudiis fácias pérfrui sempitérnis.}
    \e{O God, Whose Son hath humbled Himself, and Who hast through Him raised up the whole world, grant to thy faithful people everlasting joy; and as Thou hast delivered them from the bitter pains of eternal death, make them to be glad for ever in thy presence.}
    \l{Per Dóminum nostrum Iesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.}
    \e{Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.}
    \amen
  }
}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "vespers"
%%% End:
